%; whizzy -advi "advi -html Start-Document -edit" -mkfile ls

\documentclass{article}
\usepackage{subfiles}
\usepackage{hyperref}

\csname WhizzyShowMasterfalse\endcsname
  
\begin{document}

\section {This is the {\em master} file}
   
This is a test with the
\href
{https://ctan.tetaneutral.net/macros/latex/contrib/subfiles/subfiles.pdf}
{subfiles} \LaTeX\ package and document style.

This is an aternative to \texttt{\string\includes} that allows
a document to be split in several files and latexed either all at once or
file-by-file.  See the the
\href
{https://ctan.tetaneutral.net/macros/latex/contrib/subfiles/subfiles.pdf}
{package documentation} for details.


This text  is the {\bf master} file \verb"main.tex".   
It has two two subfiles \verb"first.tex" and \verb"second.tex"
in the same directory

This is the fist section that of the main feil.  Move to next section
for file inclusion.

\section {A master section, loading the first slave} 

The rest of this section is input from \texttt{first.tex}.
The rest of this section ends with:
\begin{verbatim}
\subfile {first}   
\end{verbatim}
\subfile {first}   


 
\section {Another master section, loading the first slave}

The next example is a few sections from an included file (hence the rest of
the slice follows on the next page). Note that, even though slices are by
section the included file, which contained several section is not sliced.
You may click on one of the included sections, which will move the pointer
and pass contol to the included file. Then, only one section of the
included file will be sliced. The sliced is delimited with lines: 
$$
\centerline{\hrulefill\space \lower 0.5ex \hbox{\bf Mastered by main.tex}
\hrulefill}
$$
on which you click to return to the master. 

The rest of this section ends with:
\begin{verbatim}
\subfile {second}
\end{verbatim}
\subfile {second}

\section {Last section, from master}

You may also input a file from a subdirectory: 
The rest of this section ends with:
\begin{verbatim}
\subfile {subdir/subdirfile}
\end{verbatim}
\subfile {subdir/subdirfile}

\section{The end}
The document ends here. 


\end{document}

