%; whizzy   document

% The previous three lines are possible configuration for whizzytex. 
% They may not be there, then, the default will be used. 
% If not present the Time-stamp line will be added if necessary. 

\documentclass[12pt]{article}

\topmargin 0em
\headsep 0in
\headheight 0in
\evensidemargin 0em
\oddsidemargin 0em
\usepackage {fullpage}

\usepackage {color}
\usepackage {hyperref}
\input {version}

\let \lst \verb

\def \whizzy {{Whizzy\kern -0.3ex\raise 0.2ex\hbox{\let \@\relax\TeX}}}
\def \Whizzy{\textbf {\textcolor {blue}{\whizzy}}}
\def \instruction #1{\par\medskip \noindent$\Rightarrow$ {\em #1}}

%% The following lines are to help HeVeA make  the HTML version of the manual

%HEVEA \def \instruction #1{}
%HEVEA\def \whizzy{{Whizzy{\TeX}}}
%HEVEA\renewcommand{\@bodyargs}{TEXT=black BGCOLOR=white}

\begin{document}
\pagestyle {empty}

\author {Didier R{\'e}my}
\date {Version {\version}, \today}
\title {
{\huge \Whizzy\footnote{Whizzytex is free software, 
Copyright \copyright 2001, 2002 INRIA
and distributed under the GNU General Public License
(See the COPYING file enclosed with the distribution).}}
\\[1em]
{\em An {\bfseries Emacs} minor-mode \\
for {\bfseries incremental viewing of} \\ 
{\bfseries {\LaTeX} documents}}
}

\maketitle   


\begin{abstract}
\def \B{\textbf}
{\whizzy}
% \footnote {Standing for {\em {\B W}hat {\B i} {\B z}ee {\B i}z what
% {\B y}ou {\expandafter \B \TeX}}}
is an Emacs minor mode for incrementally
viewing {\LaTeX} documents that you are editing.
%
It works under Unix with {\tt gv} and {\tt xdvi} viewers, but 
the \href{http://pauillac.inria.fr/advi/}{Active-DVI} viewer will
provide much better visual effects and offer more functionalities.
\end{abstract}

\section {Installation}

To use whizzytex, you need {\tt Emacs} or {\tt XEmacs}, {\tt latex2e}, and
{\tt bash} installed, and at least one DVI or Postscript previewer, such as
{\tt advi}, {\tt xdvi}, or {\tt dvips} combined with {\tt gv}. 

{\whizzy} has been developed under Linux Redhat 7.2 but has not been
extensively tested on other platforms. While {\LaTeX} and Emacs are quite
portable, you may run into compatibility problem with the bash
shell-script. 

Get the source {\tt whizzytex-\version.tgz} 
from the \href{http://pauillac.inria.fr/whizzytex}{distribution}, 
uncompress and untar it in some working directory, as follows:
\begin{quote}
\begin{tt}
gunzip whizzytex-\version.tgz\\
tar -xvf whizztex-\version.tar\\
cd whizzytex-\version
\end{tt}
\end{quote}
Then, the installation can be automatic (default or customized), or manual.

\subsection {Automatic installation}

By default, all files will be installed in the directory
\lst"/usr/local/lib/whizzytex/" and Emacs-lisp code will not be 
byte-compiled. Then, just type: 
\begin{quote}
\begin{verbatim}
make all
\end{verbatim}
\end{quote}
This will create a \lst"Makefile.config" file (only if nonexistent)
and then process configured version of the files. 

To customize the installation, you  either edit and modify 
\lst"Makefile.config", manually; you can also run 
\begin{quote}
\begin{verbatim}
make config
\end{verbatim}
\end{quote}
which will prompt for choices and override the default
\lst"Makefile.config" (which you can again edit to adjust choices). 
Check the output \lst"Makefile.config" and edit manually if necessary. 
At the end of customization, run \lst"make all" again. 

Finally, to install files to their destination, type:
\begin{quote}
\begin{verbatim}
make install
\end{verbatim}
\end{quote}

\subsection {Manual installation}

Since {\whizzy} only need three files to run, installation can also be done
manually:
\begin {itemize}

\item[]\hspace{-2em}{\tt whizzytex.el}

This could be installed in a directory visible by Emacs, but does not need
to, since you can always use the full path when you load it or declare
autoload. 

No default location.

\item[]\hspace {-2em}{\tt whizzytex}

This file is a bash-shell script that should be executable.  There is not
reason to have it visible from the executable path, since it should not be
used but with {\whizzy}.

The variable {\tt whizzytex-command-name} defined in {\tt whizzytex.el} 
contains its full path (or just its name if visible from the executable
path). 

Default value is \lst"/usr/local/lib/whizzytex/bin/whizzytex"

\item[]\hspace{-2em}{\tt whizzytex.sty}

This file are {\tt latex2e} macros. There is no reason to put this visible
from {\LaTeX} path, since it should not be used but with {\whizzy}.

Variable variable {\tt PACKAGE} defined in {\tt whizzytex} 
the full path (or just the name if the path is visible from {\LaTeX}. 

Default value is \lst"/usr/local/lib/whizzytex/latex/whizzytex.sty"

\end {itemize}

\subsection {Automatic upgrading}

For convenience, the distribution also offers a facility to download
and upgrade new versions of {\whizzy} (this requires \lst"wget" installed). 
If automatic upgrading does not work, just do it manually.

All operations should be performed in the {\whizzy} top directory, {\em
i.e.} where you untar whizzytex for the first time. We assume that have 
created a link: 
\begin{quote}
\begin{tt}
ln -s whizzytex-\version\space whizzytex
\end{tt}
\end{quote}
(which will be maintained as a link the current version by the manager). 
Alternatively, you can  use {\tt whizzytex-\version} in place of {\tt
whizzytex} below. The main commands are:
\begin{quote}
\begin{verbatim}
make -f whizzytex/Manager upgrade 
make -f whizzytex/Manager install
\end{verbatim}
\end{quote}
The command \lst"upgrade" will successively download the newest version,
unpack it, copy the configuration of the current version to the newest
version, and bring the newest version up-to-date. The command \lst"install"
will install files of the newest version. 

The following command will (re-)install and old version:
\begin{quote}
\begin{verbatim}
make VERSION=<version> download downgrade install
\end{verbatim}
\end{quote}

\section {Using {\whizzy}} 

\subsection {Loading {\tt whizzytex.el}}

Maybe, {\tt whizzytex} is already installed on your (X)Emacs system, by trying
\begin{quote}
\begin{verbatim}
ESC x whizzytex-mode
\end{verbatim}
\end{quote}
(all emacs-commands of the manual should followed by the {\sc Return} key).
If this works, skip this section.
Otherwise, you should first load the library \lst"whizzytex.el" or, better,
declare it autoload. To do this permanently, include the following
declaration in your Emacs startup file (probably is \lst"~/.emacs").
\begin{quote}\small
\begin{verbatim}
(autoload 'whizzytex-mode
    "/usr/local/bin/whizzytex/lisp/whizzytex.el" 
    "WhizzyTeX, a minor-mode WYSIWIG environment for LaTeX" t)
\end{verbatim}
\end{quote}
(where  \lst"/usr/local/bin/whizzytex/lisp/whizzytex.el" is the 
exact location of \lst"whizzytex.el", which depends on your installation.)
If  \lst"whizzytex.el" in your (X)Emacs {\tt load-path}, or if you have 
adjusted this variable appropriately, you can simply write:
\begin{quote}\small
\begin{verbatim}
(autoload 'whizzytex-mode "whizzytex" 
    "WhizzyTeX, a minor-mode WYSIWIG environment for LaTeX" t)
\end{verbatim}
\end{quote}

\subsection {Quick start} 

{\whizzy} runs as a minor mode of Emacs to be launched on a {\LaTeX}
Emacs buffer. To start it, type:
\begin{quote}
\begin{verbatim}
Esc x whizzytex-mode RET
\end{verbatim}
\end{quote}
By default, this should add new bindings so that you can later turn mode
on and off with key strokes {\tt C-c C-w}. This will also add a new menu
{\tt Whizzy} in the menu bar call ``the'' menu below.

When {\tt whizzytex-mode} is started for the first time on a new buffer, it
attempts to configure buffer local variables automatically by examining
the content of file, and using default values of global bindings.

You may customize default settings globally by running appropriate
hooks or locally by inserting appropriate comments in the source file ---see
the manual below. 

You may also change the settings interactively using the menu, or tell
whizzytex-mode to prompt the user for local customization by passing prefix
argument 4 (using, for instance, key sequence \lst"C-u C-c C-w"). 


\subsection {Editing}

Once {\tt whizzytex-mode} is on, just type in as usual.  {\whizzy} should work
transparently, refreshing the presentation as you type or move into your
{\LaTeX} buffer. 

Additionally, a gray overlay is put outside of the current slice (the {\em
slice} is the region of your buffer under focus, which is automatically
determined by {\whizzy}). When a {\LaTeX} error occurs and it can be
localized in the source buffer, a yellow overlay also is put on the region
around the error, and removed when the error is fixed.

Furthermore, when an error is persistent for several slices or some amount
of time, the interaction-buffer will pop up with the error log
(this option can be toggled on and off with the {\tt Auto interaction} menu
entry).  

\subsection {Recovering from errors}

{\whizzy} makes a good attempt at doing everything automatically. 
However, there remains situations where the user need to understand 
{\whizzy} ---when {\whizzy} does not seem to understand the user anymore. 

For that purpose, {\whizzy} report processing and error messages
in its interaction window. Thus the first help for debugging
is always to look at interaction window (buffer 
\lst"*filename.tex*" (where \lst"filename" stands for the name of the file
associated with the main buffer in case several files are composing your
document). 

This window will pop up and down automatically when an error persists or
disappear. For debugging, you may switch this {\tt Auto interaction} mode
off so as to see the buffer permanently. You may also set the mark 
to prevent the region between \lst"(point-min)" and \lst"(mark)" to be
erased automatically (as long as the buffer is visible). 

The {\tt Log...} menu entry can be used to view log files of
last actions performed by whizzytex. 

\subsection {Error during initialization}

The most delicate part of {\whizzy} is when starting {\tt whizzytex-mode},
and usually for the first time in a new buffer, since at that time all kinds
of initialization errors may occur (in addition to {\LaTeX} errors. 

{\whizzy} will attempt to identify the error and report appropriate messages
in the interaction buffer. (In case an error occurs ---or nothing happens---
always have a look at the interaction buffer, even if it did not prompted
automatically.)

Here are a description of errors during initialization mostly in
chronological order. 

\paragraph {Emacs fails during setup}

This is all under Emacs, so easily under control.
Normally, Emacs should report error messages. See the documentation for
explanations. 

In case, of uncaught fatal errors, you may
\verb"Esc X toggle-debug-on-error" to get help from Emacs, and try to fix
the problem. 

\paragraph {Emacs cannot find whizzytex}

This should typically be an installation problem, where the variable
\lst"whizzytex-command-name" is erroneous (maybe you need to give the full
path). Try to evaluate \verb"(shell-command whizzytex-command-name)" in the
minibuffer, which of course should fail, but only after the command has been
reached.

\paragraph {Whizzytex cannot build a format}

Then {\whizzy} will refuse to start.  This is most probably an error in your
macros.  Try to compile {\LaTeX} your file.  
Also try to erase all auxiliary files, since if those are erroneous, they
may disturb the creation of the format (which loads some of the auxiliary
files). 

There still could be some bad interaction between your macros and {\whizzy}
macros, but this is very rare.

\paragraph {Whizzytex cannot launch the previewer}

Usually, this is because whizzytex received wrong previewer parameter.  See
the command echoed in the interaction buffer or try to evaluate
\lst"(whizzy-get whizzytex-view-mode)".

\paragraph {Other errors}

There are two remaining problems that could happen at launch time, but that
are not particular to launch time: {\whizzy} could not recompiled the whole 
document or the current slice. However, these should not be fatal. 
In the former case, whizzytex will proceed, ignoring the whole document 
(or using the slice instead if you are in duplex mode). In the later case,
whizzytex will replace the slice by an empty slice ---and print a welcoming
document, as if you launch {\whizzy} outside of any slice. 

\subsection {Errors during normal edition}

After initialization time, {\whizzy} will keep recompiles slices as you
type or move, but also recompiles the format and the whole document when you
save a file. Each of this step may failed, but this should not be fatal, and
Emacs should report the error, possible pop up the interaction window, and
continue. 

\paragraph {Whizzytex fails on the current slice}

This should not be considered as an error, it {\bf must} happen during
edition. In particular, {\whizzy} is not much aware of {\LaTeX} and could
very well slice in the middle of the typesetting of an environment or a
macro command. This should not matter, since the erroneous slice will be
ignore temporarily until the slice is correct again.

\paragraph {Whizzytex keeps failing on the current slice}

The slice can also be erroneous because the Emacs did not correctly inferred
where to insert the cursor, which may slice erroneous, although what you
typed is correct. Hopefully, this will not occur too often, and disappear as
you move the point. It should also disappear if you switch off both {\tt
Point visible} and {\tt Page to Point} options, which is actually a good
thing to do when you suspect some misbehavior.  This will make WhizzyTeX
more robust, but less powerful and more boring.

\paragraph {Whizzytex does not seem to slice at all}

The interaction window does not produce any output. 
Try to move in the slice, or to another slice. 

If nothing happens, check the interaction
window, to see if it did attempt to recompile the slice.
If nothing happens in the interaction window, check for Emacs messages
(in the \lst"*Messages*" buffer). You may also check for the presence 
(and content) of the slice by visiting 
\lst"_whizzy_filename.tex" or
\begin{quote}
\begin{verbatim}
_whizzy_filename/input/_whizzy_name.new
\end{verbatim}
\end{quote}
If neither file exists, it means that Emacs did
not succeed to slice, which you may force by evaluating
\lst"(whizzy-observe-changes t)". 
This can be run in even if {\tt whizzytex-mode} is suspended, which may
avoid automatic processing of slices, and their erasure.

If the slice is present, you may try to compile it by hand (outside of
Emacs) with 
\begin{quote}
\begin{verbatim}
latex '&filename' _whizzy_filename.tex
\end{verbatim}
\end{quote}
and see if it succeeds. 


\paragraph {Reformatting failed}

Formatting errors are fatal during initialization, but accepted once
initialized. In case of an error during reformatting, {\whizzy} will ignore
the error and continue with the old format.  This means that new macros may
be ignored leading to further slicing errors. When rebuilding the format
failed, the mode-line string will display the suffix \lst"FMT" until the
error is fixed.  See the interaction buffer or select \lst"format" from the
\lst"log..." menu entry).

You may also force reformatting by typing the \lst"reformat" command
in the interaction buffer. 


\paragraph {Whizzytex cannot process the whole document}

This is very likely a problem with you document, so try to {\LaTeX} it 
first. There is a small possibility of strange interaction between
your macros and {\whizzy} package. Try to turn options 
{\tt Page to Point} and {\tt Point visible} off and retry. 
This will make {\whizzy} more robust (but also less powerful and more
boring). 


\section {\label{manual}Manual} 

This section describes how to use and parameterize {\whizzy}. 

Since, most of the documentation is available online in Emacs from the
\lst"Help" entry of the menu by following hyperlinks, we keep this section
to a minimal, so as to avoid redundancy. 

Alternatively, you can type
\begin{quote}
\begin{verbatim}
ESC x describe-function RET whizzytex-mode
\end{verbatim}
\end{quote}
(In XEmacs, you may need to use
\begin{quote}
\begin{verbatim}
ESC x hyper-describe-function RET whizzytex-mode
\end{verbatim}
\end{quote}
instead of \lst"describe-function" to see hyper-links.)

Section \ref{configuration}, \ref{modes} and
\ref{types} are also available as online help. 

\subsection {Configuration} 

\label{configuration}

\subsubsection {Emacs global configuration}

\label {Emacs-configuration}

See Emacs help for \lst"whizzy-default-bindings" and
\lst"whizzytex-mode-hook" for list of bindings.

See help for \lst"whizzytex-mode" a  list of user-configurable variables, 
to which you may give default values in your Emacs startup file that will be
used instead of {\whizzy} own default values. 

\subsubsection {File-based configuration}

\label {File-configuration}

A configuration line is one that starts with regexp prefix ``\lst"^%; +"''
followed by a configuration keyword.  If two configuration lines have the same
keyword, only the first one is considered. The argument of a configuration
line is the rest of the line stripped of its white space.

The keywords are:
\def \arg#1{$\langle\texttt {#1}\rangle$}
\def \opt#1{[ #1 ]}
\begin{description}
\item[whizzy-master]\arg {master}
\\
This only makes sense for a file loaded by a {\em master} file. 
\arg{master} is the relative or full name of the
master file. Optional surrounding quotes (character \lst$"$) are
stripped off, so that \lst$"foo.tex"$ and \lst"foo.tex" are equivalent.

\item[whizzy] 
\opt{\arg{slicing}} 
\opt{\arg{viewer} \opt{\arg{command}}}
\opt{\texttt{-pre} \arg{make}}
\opt{\texttt{-duplex}}
\\[1em]
All arguments are optional, but if present they must appear in order:
\begin{description}
\item[\arg{slicing}]\indent\\ 
is determines the way the document is sliced
(see section~\ref{modes}).

\item[\arg{viewer}]\indent\\
is the type of viewer and can only be one of 
\lst"-advi", \lst"-xdvi", or \lst"-ps" (see section~\ref{types})

\item[\arg{command}]\indent\\
is optional and is the command used to call the viewer
(of course, it should agree with \arg{viewer}). 

\item[\texttt{-pre} \arg{make}]\indent\\
tells {\whizzy} to use the command \arg{make} to preprocess
the slice. 
The command \arg{make} will receive one argument
\texttt{basename.new} and should produce \texttt{basename.tex}. 
By default, \lst"mv" is simply used.

The Unix \lst"make"  can itself be used as a preprocessor (with an
appropriate \lst"Makefile").  However, one may have to work around
\lst"make"'s notion of time (using FORCE), which is usually too rough. 
This is safe, since {\whizzy} tests itself for needed recompilations.

\item[{\bf {\tt -fmt FORMAT}}]\indent

By default {\whizzy} build a specialized format of the \lst"latex" document
format by loading macros of your document.  The above option will use
\lst"FORMAT" instead of \lst"latex" as the initial format. For instance,
\lst"hugelatex" could be used (depending on your {\LaTeX} configuration) to
build a larger format to process huge files.

\item[\texttt{-duplex}]\indent\\
tells {\whizzy} to launch another window with the whole document (which is
recompiled every time the source buffer is saved).

With \lst"-advi" previewers, both views  communicate with Emacs and can be
used to navigate through source buffers and positions.
\end{description}

For instance, a typical configuration line will be:
\begin{verbatim}
   %; whizzy -dvi xdvi
\end{verbatim}
tells whizzytex not to take the whole document as a slice
use a \lst"dvi" 
previewer called with the command \lst"xdvi"

More evolved configuration lines are:
\begin{verbatim}
   %; whizzy section -advi dview -whizzy -pre make -duplex
   %; whizzy section -advi "dview -whizzy" -pre make -duplex
\end{verbatim}
These two lines are equivalent, since Emacs will remove outer quotes in
option arguments. They both tell whizzytex to be be launch in \lst"section"
mode and use a \lst"dvi" previewer called with the command 
\lst"dview -whizzy" 
(a shell-script to be defined), use \lst"make" to preprocess files
and provide \lst"duplex" views.

\item[whizzy-paragraph] \texttt{regexp}\\
This sets the Emacs variable \lst"whizzy-paragraph" to \texttt{regexp}.
\end{description} 


\subsection {Modes} 
\label {modes}

{\whizzy} recognizes three modes \lst"slide", \lst"section", and \lst"document". 
The mode determines the slice of the document being displayed and how
frequently updates occurs. 
\begin{description}

\item [slide]

The mode \lst"slide"  is mainly used for documents of the class seminar. 
In slide mode, the slide is the text between two \lst"\begin {slide}"
comments (thus,  the text between two slides is displayed after the
preceding slide).  

In slice modes, overlays are ignored {\em i.e.} all overlays all displayed in
the same slide, unless a command
\lst"\overlay {"$n$\lst"}" occurs on the left of the point, on the same line
(if several commands are on the same line, the 
right-most one is taken), in which case only layers $p \le n$ are displayed.

\item [section]
In \lst"section" mode, the slice of text is the current chapter, section.

\item [subsection]
As \lst"section" but also slice at subsections. 

\item [paragraph]
The \lst"paragraph" mode is a variation on section mode where, the separator
\lst"whizzy-paragraph" is defined by the user (set to two empty lines by
default) instead of using \lst"\section"  and \lst"\subsection" commands. 
subsection.

\item [document]
The \lst"document" take the region between \lst"\begin{document}"
and \lst"\end"\lst"{document}" as the slice. 

\item [none]
In \lst"none" slicing mode, there is no sectioning unit at all and
the whole document is recompiled altogether. 
Currently, pages are not turned to point and the 
cursor is not shown in \lst"document" mode, because full documents are not
sliced. (A slicing document mode could be obtained by working in paragraph
mode, with an appropriate regexp.)

\end{description}

\subsection {Viewer types}
\label {types}

See help for \lst"whizzy-viewers".

The previewer types can have three possible values:
{\tt -ps}, {\tt -dvi}, or {\tt -advi}. 

The previewer type should agree with the previewer command in several ways:
\begin {itemize}

\item
They tell whizzytex whether to use signal {\tt SIGHUP} (with {\tt -ps})
or {\tt SIGUSR1}  ({\tt -dvi} and {\tt -advi}) to tell the previewer to
reload the file

In particular, if you write a front-hand shell-script \lst"viewer" to the
call previewer,  and want to use \lst"viewer" as the previewer, you should 
arrange for \lst"viewer" to understand these signals (and forward them to the
previewer). The simplest way is to hand your script with an exec command
calling the \lst"gv", \lst"dvi" or \lst"advi".



\item
They tell whizzytex whether to process the slice to 
Postscript (with {\tt -ps}) or to DVI format (with {\tt -dvi} and {\tt -advi}). 

\item
Moreover, {\tt -advi} requires the previewer to 
recognize additional \lst"\special" commands, in particular
source line information of the form: 
\begin{quote}
\begin{verbatim}
#line 610, 615 <<recog>><<nizes>> manual
\end{verbatim}
\end{quote}

\end {itemize}
Then, the previewer command is the command to call the previewer.  This
string will be passed as such to the {\whizzy} shell-script. Note that the
name of the \lst"dvi" or postscript file will be appended to the previewer
command.


\subsection {Watching other files}

{\whizzy} is designed to watch other files and not just the slice saved by
Emacs. In fact, it watches any file dropped in the pool directory. 
For instance, 
if your source file uses images, you can just change the image and
drop the new version in the pool. Then {\whizzy} will pick the new version,
move it to the working directory and recompile a new slice. Be aware of name
clashes: if you drop a file in the pool, it will automatically be move to
the working directory with the same name, overriding any file of the same
name sitting there. 

By default, {\whizzy} send itself a \lst"kill -STOP" signal when it finds
nothing in the pool (this allows you to let {\whizzy} inactive for hours
without consuming any CPU. Correspondingly, when Emacs drops a new file in
the pool, it sends a \lst"kill -CONT" signal. If you want to watch other
files and not have to deal with waking up {\whizzy} you can pass the extra
option \lst"-nonstop" to the \lst"whizzytex" command-line; then it will sleep
instead of stopping itself, and will periodically watch for news files 
in the pool without you need to signal it to continue. 
In case you want to keep with the stop-mode and wake it up explicitly, the
process number is stored in the file \lst"._whizzy_$BASICNAME.tex". %$

\subsection {Frequency of recompilation} 

To obtain maximum {\whizzy} effect, a new slice should be save after any
edition changed or any displacement that outside of the current slice.
However, to avoid overloading the machine with useless and annoying
refreshments, some compromise is made, depending on Emacs several
parameters: edition {\em v.s.} move Emacs last commands, 
successful {\em v.s.} erroneous last slice, and the duration of last slice
recompilation. This usually works well. However, different behavior may wish
to be obtained in different situations. For instance, when editing on a
lab-top, one may wish to save batteries by keeping the load rather low, hence
not using the full power of the processor. Conversely, one may wish 
{\whizzy} to be as responsive as possible. There is an function
\lst"whizzy-load-factor" that control a variable of the same name, which can
be used to adjust the responsiveness (by increasing or decreasing the
load-factor). This simply adds extra delays between slicing. 

The format is automatically recompiled at the beginning of each session, and
whenever the buffer containing the file is saved. That is, to load new
packages or define new global macros (before the \lst"\begin{document}"), it
suffices to save the current file.

\subsection {Cross-references, page and section numbers} 

The slice is always recompiled with the \lst".aux" file of the whole
document.  In paragraph mode, cross references and section numbers are 
recompiled whenever the buffer itself is saved (manually). 
This recompilation operates in batch and concurrently with the recompilation
of slices, so it may take several slices before the new counters or
references are adjusted.

(The recompilation of the whole document is off in slide mode.) 


\section {Viewers}

\subsection {Viewing with Active-Dvi}

\href{http://pauillac.inria.fr/advi/}{Active-DVI} is a DVI previewer with
several additional features. 
In particular, it recognizes extra specials, some of which are particular 
useful for whizzytex that allows a two way communication between 
the source Emacs buffer and the previewer: 
\begin {itemize}
\item
The previewer will automatically turn pages for you, as you are editing. 
This is done by telling Emacs to save the current position in the slice. 
Then, the recompilation of the slice will include the current position 
as an hyperref location \lst"Start-Document" whenever possible. 
Then, just tell \lst"Active-Dvi" to automatically jump at this location
when it opens/reloads the file. 

\item
Conversely, \lst"Active-Dvi" can dump source file positions on clicks, 
when available, that is forwarded to Emacs so that it can move to the
corresponding line.

To enjoy this feature, the option \lst"-advi" should be used instead of
\lst"-dvi". This will produce extra information (such as source line
numbers) using \lst"\special" that most DVI previewers do not recognize
and may complain about.

\end {itemize}
For instance, your could then set
Hence, my Emacs setup executes
\begin{quote}
\begin{verbatim}
(set-default 'whizzy-viewers
  (("-advi" "advi -html Start-Document") 
   ("-dvi" "xdvi") ("-ps" "gv")))))
\end{verbatim}
\end{quote}

\subsection{Defining your own previewer}

To use your own command as a previewer, you must choose either type
\lst"-dvi" or \lst"-ps" . In particular, your previewer should 
accept \lst"SIGUSR1" (for \lst"-dvi") signal or \lst"SIGHUP" (for \lst"-ps") 
signal and respond by reloading the file.

\subsection{Viewing with acroread}

This does not work because they is no simple way to tell
\lst"acroread" to reload its file in batch. 

\section {Implementation}

In short, {\sc \whizzy} is selecting a small slice of the document that 
you are editing around the cursor (according to the selected mode) 
and redisplay the slice incrementally as it changes through edition. 
\begin {itemize}

\item {\bf Emacs is watching you} typing and moving in the 
Emacs buffer attached to the {\LaTeX} source file that your editing and keeps
saving the current slice (current slide, section, or subsection, according
to the mode).

\item {\bf A shell-script daemon}
keeps recompiling whenever a new slice (or other files) are produced, and if
recompilation succeeds, tels the previewer to updates the display of the slice.

\item {\bf A few {\LaTeX} macros} allow to build a specialized
format with all macro loaded, which considerably speed up the time for
slicing. Additionally, the slice is a bit instrumented to show the cursor,
and includes specials that allows back-pointing from the DVI file into the
Emacs buffer.

\end {itemize}
The rest of this section briefly describe these three parts\footnote {This
section is not quite up-to-date, hence it puts emphasis on the original
design, but several aspects have changed significantly since the first
version. Implementation of more recent features is thus omitted.}, and
their interactions.


\subsection {Emacs code}

The main trick is to use \lst"post-command-hook" to make Emacs watch
changes.  It uses \lst"buffer-modified-tick" to tell if any editing has
actually occurred, and compare the point position with the (remembered)
position of the region being displayed to see if saving should occur.  It
uses \lst"sit-for" to delay slicing until at least the time of slice
computation has ellapsed since last saving, a significant number of editing
changes has occurred, or iddleness.

{\whizzy} can also display the cursor position, in which case slices are
also recomputed when the cursor moves, but with lower priority.

\subsection {{\LaTeX} code}

The main TeX trick is to build a format specialized to the current
document  so as to avoid reloading the
whole macros at each compilation. This is (almost\footnote{{\tt
$\backslash$begin\{document\}} should be typed as such without any white
white space}) entirely transparent, that is, the source file does not have
to understand this trick.

This is implemented by redefining \lst"\documentclass" which in turn
redefines \lst"\document" to execute \lst"\dump" (after redefining
\lst"\document" to its old value and \lst"\documentclass" so that it skips
everything till \lst"\document"). This is robust ---and seems 
to work with rather complex macros. 

The specialized format can be used in two modes: by default it expects a
full document: it them dumps counters at sectioning commands (chapters, 
sections, and subsections). This is useful to correctly 
numberred sections and pages on slices. 

There are also a a few other used to get more advanced behavior, especially
to dump source line numbers and file names so that the previewer can
transform clicks into source file positions. 

When building the format, {\whizzy} also look for a local file of name
\lst"whizzy.sty", which if existing is loaded at the end of the macros. 
This may be used to add other macros in {whizzy} mode, {\em e.g.} 
some {\TeX} environments may be redefined to changed they type setting,
according to whether the current line is inside or outside the environment. 
(We have written such an extension for an exercise package that sends the
answers at the end in an appendix, unless the cursor is inside the answer,
in which case the answer is in-lined.)

\subsection {Bash code}

There is no real trick there. This is a shell-script watching the pool
(a directory where slices and other new version of files must be dropped). 
It them recompiles a slice and wait for input (in stdin). 
It recognizes a few one-line commands as input {\tt reformat}, {\tt
dupplex}, and by default just watch for the presence of a new slice. 
It recompiles the format file (and the page and section number, but in batch
mode) whenever the source file (its Unix date) has changed  and 
recompiles the slice whenever it is present (since {\whizzy} renames ---hence
removes--- the slice before processing it).

If the file has been recompiled successfully, it triggers the previewer
(ghostscript or xdvi) so that it rereads the dvi or ps file. Otherwise, it
processes the {\TeX} log file and tries to locate the error. It then sends part
of the log file with annotations to the \lst"*TeX-shell*" buffer from which
Emacs has been {\whizzy}, so that Emacs can report the error. 

\subsection {Interaction between the components} 

The control is normally done by Emacs, which launches and kills the Unix
daemon. Quitting the previewer should be noticed by the daemon, which tells
Emacs to turn mode off before exiting. 

Muliple {\whizzy} running on the same file would certainly raise racing
conditions between files and would not make sense. 
For that purpose, the daemon pid is saved in a file and {\whizzy}
will kill any old {\whizzy} process on startup. 

\section {Whizzy dream}

When used together with Active-DVI, {\whizzy} could be made much mode
powerful. In particular, it would be quite easy to to lift {\whizzy} from an
incremental viewer to an assistant editor.

Active-DVI could easily provide a notion of active boxes.
These would be recognized by
\lst"\special" annotations preceeding  boxes. 
Active boxes would be autoraise on focus and could be moved or resized with
the mouse. Rather than displaying actions on screen, which would be unaware
of {\TeX} position stategies, actions should rather be reported in stdout,
as is already done for positions. 

Active-DVI actions could then easily be interepreted by {\whizzy} by
adjusting or inserting the correct vertical or horizontal dimension around
active objects, and processed in the next slice. Thanks to the short
incremental loop, this would (almost) appear as if actions where executed by
Active-DVI.



\end{document}

- Clicks

- Say that cursor does not always appear.
% LocalWords:  whizzy advi html whizzytex HeVeA BGCOLOR Didier INRIA ee ou ing
% LocalWords:  ghostview xdvi DVI XEmacs Redhat tgz untar gunzip xf usr config
% LocalWords:  Esc tex minibuffer FMT online dvi dview args autoload ps pre mv
% LocalWords:  BASENAME hyperref gv HUP pid dvips xvf cd wget WhizzyTeX WYSIWIG
% LocalWords:  RET preprocess basename fmt hugelatex BASICFILENAME SIGHUP recog
% LocalWords:  SIGUSR nizes acroread localmacros ghostscript
