%; whizzy -advi "advi -html Start-Document -edit" -mkfile ls

\documentclass{article}

\csname WhizzyShowMasterfalse\endcsname
  
\begin{document}

\section {This is the {\em master} file}
   
This is a test with multiple files.  
This  is the {\bf master} file \verb"main.tex".   
It has two slaves files \verb"first.tex" and \verb"second.tex". 
This is the fist section that does not input any file.
Move to next section for file inclusion.

\section {A master section, loading the first slave} 

The rest of this section is input from \texttt{first.tex}.
The rest of this section ends with:
\begin{verbatim}
\input {first}   
\end{verbatim}
\input {first}   


 
\section {Another master section, loading the first slave}

The next example is a few sections from an included file (hence the rest of
the slice follows on the next page). Note that, even though slices are by
section the included file, which contained several section is not sliced.
You may click on one of the included sections, which will move the pointer
and pass contol to the included file. Then, only one section of the
included file will be sliced. The sliced is delimited with lines: 
$$
\centerline{\hrulefill\space \lower 0.5ex \hbox{\bf Mastered by main.tex}
\hrulefill}
$$
on which you click to return to the master. 

The rest of this section ends with:
\begin{verbatim}
\include {second}
\end{verbatim}
\include {second}

\section {Last section, from master}

You may also input a file from a subdirectory: 
The rest of this section ends with:
\begin{verbatim}
\input {subdir/subfile}
\end{verbatim}
\input {subdir/subfile}

\section{The end}
The document ends here. 


\end{document}

